### 2022-02-07 input/4_user_study_and_methods.tex

% Gender-inclusive: Relevant throughout the document but frequently occuring here:
% - [ ]  write gender-inclusive: do not use he or she but use they instead: "the participant was asking..., he liked ... → they liked (in singular)
% check for the following:
% - [ ]  Have you used “man” or “men” or words containing them to refer to people who may not be men?
% - [ ]  Have you used “he,” “him,” “his,” or “himself” to refer to people who may not be men?
% - [ ]  If you have mentioned someone’s sex or gender, was it necessary to do so?
% - [ ]  Do you use any occupational (or other) stereotypes?
% - [ ]  Do you provide the same kinds of information and descriptions when writing about people of different genders?

\section{User Study}

\section{Methods}