### 2022-09-04 input/5_results.tex

\section{Results}

\subsection{Human-Computer Collaboration for Human Augmentation}
\missingfigure{A: Boxplot of questionnaire scores, are they significantly better than 0? Are they significantly better then after the training? -> frame the training the same as the experiment, B: Sentiment Word Clouds}

% system usability score

% sentiment analysis ttest the counts?

% Did they get faster? Calculate time between hand motion onset and button press. If yes, evidence that there was acceptance for the help/augmentation the system provided

\subsection{The BCI}

\missingfigure{!: grand average ERP at Cz comparing idle vs. movement intention with scalp map of voltage across all channels in the last [-100 0] interval of both idle and pre-move; B: What did the windows mean features look like. Boxplot with idle vs. pre-move and single observations}

% ttest of amplitudes in last time window

% classification accuracy

\missingfigure{confusion matrix!, AUC, ROC -> in discussion mention that we want to do that continuously and not only look at the times preceding hand movement but at all time points}