### 2023-04-28 input/7_discussion.tex

\section{Discussion}

% summary here! BLUF!

% Outlook Ideas
% - Strength of EMS could be set by the amplitude of the RP -> is there evidence for a correlation between amplitude and 'strength' of volitional action?
%[ - EEG Signal may be contaminated by eye movement, maybe a worthwhile thing to build this interaction relying solely on eye tracking information, gaze fixating on the target before moving]
%- not only physical integration but also other adpative interfaces, or control of mobile phone (see Cornelio2022 review); % say here somewhere what action augmentation is and how it differs from body and environment/outcome augmentation