### 2023-07-26 input/6_results.tex

\section{Results}

\subsection{Agency}
\missingfigure{A: Boxplot of questionnaire scores of three Blocks (block 2 grey, because not intention given) - play around with visualization of words that were used to explain control rating. B. Boxplots intentional binding (maybe integrate A and B) C. more subjective evaluation of interview / maybe world cloud / maybe sentiment analysis) }

% sentiment analysis ttest the counts?

% Did they get faster? Calculate time between hand motion onset and button press. If yes, evidence that there was acceptance for the help/augmentation the system provided

\subsection{BCI performance}

\missingfigure{!: grand average ERP at Cz comparing idle vs. movement intention with scalp map of voltage across all channels in the last [-100 0] interval of both idle and pre-move; B: What did the windows mean features look like. Boxplot with idle vs. pre-move and single observations}

% ttest of amplitudes in last time window

% classification accuracy

\missingfigure{confusion matrix!, AUC, ROC -> in discussion mention that we want to do that continuously and not only look at the times preceding hand movement but at all time points}