### 2023-07-26 input/7_discussion.tex

\section{Discussion}

% summary here! BLUF!

% Outlook Ideas
% - Strength of EMS could be set by the amplitude of the RP -> is there evidence for a correlation between amplitude and 'strength' of volitional action?
%[ - EEG Signal may be contaminated by eye movement, maybe a worthwhile thing to build this interaction relying solely on eye tracking information, gaze fixating on the target before moving]
%- not only physical integration but also other adpative interfaces, or control of mobile phone (see Cornelio2022 review); % say here somewhere what action augmentation is and how it differs from body and environment/outcome augmentation

% playful experiences: EMS

\begin{comment}
% snippets that might be useful
- a sense of agency is not an infallible reproduction of objective reality / can be quite divorced from objective reality (see Moore 2016) // --> if people still report agency, in no-intent (only EMS) condition

\end{comment}