### 2024-04-03 UIST_submission/source/input/0_abstract.tex
UIST submission

% To maintain a sense of agency (SoA), physical motor augmentation must align with the users' expectations. In experimental setups, this is often achieved using stimulus-response paradigms where the user behavior is predictable. Moving towards user augmentation in the everyday world where user behavior is less predictable, we developed a brain-computer interface (BCI) based augmentation using electrical muscle stimulation (EMS). By classifying readiness potentials from the user's electroencephalogram (EEG), our system controlled the user's movements at the time of their intent to interact. The system was able to discriminate pre-movement from idle EEG segments with an F1 score of 0.7. In a pilot study, we observed intentional binding, a compression of time between action and outcome, and a higher level of control when participants worked with the system instead of being passively moved. This is a step towards predictive interfaces that leverage the users’ bodies yet still feel natural as they align with the users’ intentions.

To maintain a sense of agency (SoA), physical motor augmentation must align with the users' intentions. In experiments, this is often achieved using stimulus-response paradigms. Here, the user behavior follows the stimulus and motor augmentation can be delivered in line with the user's motor plan. But, in the everyday world user's act on their own volition. Here, we designed a system to control the user's movements at the time of their intent to interact. A brain-computer interface (BCI) classified readiness potentials from the user's electroencephalogram (EEG). The BCI's output then gated a motor augmentation via electrical muscle stimulation (EMS). The system discriminated pre-movement from idle EEG segments with an F1 score of 0.7. In a pilot study, we found evidence for a maintained SoA. Participants mentally compressed the time between action and outcome, intentional binding, and reported a higher level of control when working with the system instead of being passively moved. This is a step towards leveraging the users’ bodies while maintaining SoA.