### 2024-10-04 submission_JNE/source/input/2_introduction.tex
page and jne submission

\section{Introduction}

Advances in hardware that augment a user's physical actions have reignited dreams of overcoming human limitations, recovering lost abilities and simplifying skill acquisition~\citep{Goto2020-mw, Kunze2017-co}. These technological advances include the miniaturization of the actuating hardware to wearable form factors and the direct sensing and stimulation capabilities of neural interfaces. Especially due to these characteristics, recent perspectives promote a change of the computing era from human-computer \textit{interaction} to \textit{integration}~\cite{Mueller2020-dl}. One key change in perspective is, that \textit{integrated} users share agency with the computing machinery to execute tasks. This is a critical distinction as \textit{integration} technologies are designed to directly influence people’s bodies, their actions, and the resulting action outcomes~\cite{Cornelio2022-aq}.

Besides body and outcome augmentation, action augmentation has been defined as the case where a ``system assists the user’s action to produce the intended outcome.''~\cite{Cornelio2022-aq}. Such action augmentations can be realized purely on a software integration level, for example by an AI pair programmer. When it is designed to happen on a hardware level, further challenges emerge, specifically due to shared agency, which in this case means handing over control of one's own body. Unfortunately, such augmented users often report dissociative experiences, frequently disrupting their sense of agency (SoA)~\citep{Gilbert2017-ze, Gilbert2019-uc}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/teaser.png}%{figures/bci_game.png}
    \caption{We propose an augmentation system that aligns with the user's agency. When participants feel the spontaneous urge to move, readiness potentials (RPs) are picked up in the user's brain signals. A brain-computer interface (BCI) then predicts the user to be in either an idle (red) or pre-movement (green) state. In the latter case, electrical muscle stimulation (EMS) is triggered and moves the user's hand. Image taken with consent from participant.}
    \label{fig:teaser}
\end{figure}

Having an SoA means experiencing control over our own voluntary actions, instead of them feeling as randomly happening to us, for a recent review see~\cite{Wen2022-bm}. It has been shown that users are more likely to feel engaged and satisfied with an interaction, and are more likely to trust a system the more they experience SoA~\citep{Berberian2012-do, Miller2007-rb}. Hence, a key challenge to drive the adoption of human action augmentation is to design for agency experience, so users feel as though they are in the ``driver's seat'' once again.

% \subsection*{Preserving Agency using Brain Signals of the Intent to (Inter-) act}
With this paper, we strive towards an `agency-aligned' augmentation system. We developed a brain-computer interface (BCI) that establishes a fast communication channel between a user's brain signals and a physical end effector, see figure~\ref{fig:teaser}. The closed-loop augmentation system (from hereon referred to simply as \textit{system}) controls the user's muscles at the time of their \textit{intent to interact}, as measured through readiness potentials (RP) manifesting in the user's electroencephalogram (EEG)~\cite{Schurger2021-vp, Schultze-Kraft2016-bx, Schultze-Kraft2021-cu}. Ultimately, such an augmentation can cue the user's movements, increase their strength, and might also preempt their action, i.e., increase their speed.

% \begin{figure}[!h]
%     \centering
%     \includegraphics[width=\columnwidth]{figures/teaser_new.png}%{figures/bci_game.png}
%     \caption{Our augmentation system: When participants feel the spontaneous urge to move, readiness potentials (RPs) are picked up in the user's brain signals. A brain-computer interface (BCI) predicts the data to be in either of two classes: Idle or pre-movement. In the latter case, electrical muscle stimulation (EMS) is triggered and the user's hand is moved. Image taken with consent from participant.}
% \end{figure}

To control the interaction in real-time, an RP-based classifier distinguished between two user states: \textit{idle}, reflecting the absence of an intent to act, and \textit{pre-movement}, indicating the presence of an intent to act. During \textit{idle}, participants were passively looking at a fixation cross. Instead, during \textit{pre-movement}, participants were instructed to voluntarily initiate a tap on a touchscreen whenever they felt the urge to do so. Previous work has indicated that the RP emerges during formation of conscious intention and is specific to voluntary action~\cite{Schultze-Kraft2020-rm, Travers2020-hf, Pares-Pujolras2019-ll}.

Upon predicting a \textit{pre-movement} state, the system augments the user's action, potentially even preceding their voluntary motor command. This augmentation moves the ring finger in accordance with the user's intention to act. We achieved the movement by leveraging electrical muscle stimulation (EMS) applied to the user's forearm flexor muscle. To get a first glimpse of how our system can maintain agency in augmented users, we conducted a user study using both quantitative and qualitative research methods. In short, we explored whether aligning the physical impact on the user's body with their intention to move, preserved their SoA.