### 2024-10-04 submission_JNE/source/input/5_user_study_and_methods.tex

\section{User study \& Methods}
In this paper, we present a prototype that uses a user's own brain signals as the control signal to a physical end effector. With a user study, we wanted to find out whether the experience of agency can be preserved during physical action augmentation when using our prototype.

We compared three conditions of agency experience. With the first two conditions, INTENTION and INVOLUNTARY we queried users at the two edges of agency experience. In INVOLUNTARY, participants had no intention to move and no control over their movement, much like in a bystander scenario. On the other hand, in INTENTION, participants were acting as they would in their day-to-day lives, fully in control and with volition. We then investigated in a third condition where our AUGMENTED prototype sits on the spectrum between INTENTION and INVOLUNTARY, in other words, whether it preserved agency and to what degree. To answer our question, we employed a mix of qualitative and quantitative methods, including a psychometric test of intentional binding, one standardized question, a qualitative (phenomenological) exit interview, and an exploratory EEG analyses of prediction errors following muscle stimulation.

% To assist readers in replicating our prototype and experiment, we provide the necessary technical details, the complete source code for the BCI and the experiment, the collected data, as well as all the analysis scripts\footnote{[anonymized]}
% https://github.com/lukasgehrke/2021-fastReach

\subsection{Participants}
Eleven participants (M = 29.9 years, SD = 4) were recruited from our local institution and through the institute's online participant pool. Participants were compensated with course credit or 12 Euro per hour of study participation. Prior to their participation, they were informed of the nature of the experiment, recording, and anonymization procedures and signed a consent form. The experiment was approved by the local ethics committee of the Department of Psychology and Ergonomics at the TU Berlin (Ethics protocol approval code: BPN\_GEH\_2\_230130). One participant had to be excluded from further data analyses due to significant deviations from the instructions in the execution of the task. Precisely, they did not initiate their movements at their own volition but rather immediately at the onset of each trial, thereby violating the task instruction of waiting 2-3 s after trial onset, see task description below for more detail.

\subsection{Apparatus}
The experimental setup, depicted in Figure~\ref{fig:setup}, comprised: (1) a 1-channel Electromyography (EMG) device, (2) a 64-channel EEG system, (3) a medically-compliant EMS device connected to two electrodes worn on the forearm, and (4) a tablet to run the experiment and collect behavioral responses.

\begin{figure}[!h]
    \centering
    \includegraphics[width=\columnwidth]{figures/setup_fixed.png}
    \caption{Experimental setup of measurement-- and input devices (image with consent from participant).}
    \label{fig:setup}
\end{figure}

\indent\textbf{(1) EMG Recording.} EMG data was recorded from 1 bipolar channel using the BrainAmp ExG amplifier (BrainProducts GmbH, Gilching, Germany). The two electrodes were placed above the flexor digitorum profundus with a reference electrode located on the wrist bone. EMG data was collected in synchrony with the EEG data through BrainProducts' BrainVision Recorder. EMG data was only recorded for the first experimental condition to obtain labels for classifier training. After that, the EMG electrodes were changed for EMS electrodes.

\indent\textbf{(2) EEG Recording.} EEG data was recorded from 64 actively amplified Ag/AgCl electrodes referenced to electrode FCz in an actiCap Snap cap using BrainAmp DC amplifiers from BrainProducts. Electrodes were placed according to the extended international 10–20 system \cite{Jasper1983-uw}. One electrode was placed under the right eye to provide additional information about eye movements (vEOG). After fitting the cap, all electrodes were filled with conductive gel to ensure proper conductivity, and electrode impedance was brought below 10k$\Omega$ where possible. EEG (and EMG) data were recorded with a sampling rate of 250 Hz.

We used LSL\footnote{https://github.com/sccn/labstreaminglayer} to make the data streams available in the network and synchronize the recordings of EEG/EMG data and an experiment marker stream that marked sections of the study procedure.

\indent\textbf{(3) Electrical Muscle Stimulation.} We actuated the ring finger via EMS, which was delivered with two electrodes attached to the participants' flexor digitorum profundus muscle. We utilized the flexor digitorum profundus since we found that we can robustly actuate it without inducing unintended motion of neighboring fingers. This finger actuation was achieved via a medically compliant battery-powered muscle stimulator (TENS/EMS Super Duo Plus, prorelax, Düren, Germany). The EMS system's output was controlled by flipping a solid state relay (silent) connected via an Arduino Uno (Arduino, Monza, Italy) to the experiment computer. The EMS was pre-calibrated by the participant to ensure a pain-free but effective stimulation and robust actuation leading to an `immediate' tap on the touchscreen after actuation. To ensure a comfortable experimental experience so that participants relaxed their arm musculature as much as possible, a custom built hand rest (support device) was placed on top of the touchscreen for participants, see figure~\ref{fig:setup}.

\indent\textbf{(4) Experiment Presentation and Collection of Behavioral Responses.} An Acer Group (Acer Inc, Taipeh, Taiwan) tablet was used to present the task to participants and record their behavioral responses. In addition to the tablet, we used an INVOLUNTARY keyboard to allow users to input their timing judgments.

\subsection{Experimental Task, Design and Procedure}
Participants performed a simple tapping task in three conditions with 75 trials each: INTENTION, INVOLUNTARY, and AUGMENTED

\begin{figure}[h]
    \centering
    \includegraphics[width=\columnwidth]{figures/task_new.pdf}
    \caption{Interaction flow depicting one trial in our touchscreen tapping task.}
    \label{fig:progression}
\end{figure}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/method.png}
    \caption{Diagram summarizing(a) signal processing and event labeling from muscle activity, (b) extraction of linear slope features from labeled brain signals, and (c) real-time application of the BCI gating the muscle stimulation hardware.}
    \label{fig:method}
\end{figure*}

\indent\textbf{INTENTION.} The task was as follows: (1) a fixation cross appeared on a tablet screen and participants were instructed to rest and wait until it disappeared; (2) they were instructed to wait for a brief moment (2 to 3s), before (3) initiating their movement and tap the screen, see figure~\ref{fig:progression}. In line with the literature on the origin of the RP generating process, they were told ``to avoid pre-planning the movement, avoid any obvious rhythm across trials, and to press when they felt the spontaneous urge to move''~\cite{Schultze-Kraft2021-cu}. (4) After the screen was tapped, a tone was played at a pseudo-random delay of 200, 350, or 500ms. Participants were now asked to estimate the delay, typing in their answers on a number pad of an attached keyboard. After confirming their answer by hitting the return key, the next trial started.

During INTENTION, participants were equipped with EMG sensors instead of EMS electrodes. Participants' reaction times in INTENTION were used to select stereotypical reaction times for the INVOLUNTARY condition. At the end of INTENTION, EMG electrodes were exchanged for EMS electrodes at the identical location on the forearm.

\indent\textbf{INVOLUNTARY.} The task structure was identical to INTENTION, however, participants were now instructed to hold and wait for the muscle stimulation to move their finger thereby eliciting the screen tap. The timing for the EMS trigger was taken by randomly choosing a time between the 5th and 95th percentile of their actual individual reaction time in the INTENTION condition.

\indent\textbf{AUGMENTED.} The task and instruction were identical to INTENTION with one additional instruction: ``you will now work \textit{with} the system''. During AUGMENTED, the muscle stimulation hardware was controlled by the BCI. The classifier was set to active after the fixation cross disappeared and until a screen tap was registered. Hence, the muscle was not stimulated at other times during a trial, so as not to interfere with participants typing in their time estimation response.

The order of the conditions was not pseudo-randomized since training data obtained in INTENTION was required for both INVOLUNTARY and AUGMENTED. Furthermore, AUGMENTED was always the last condition, thereby allowing for a prolonged interview immediately after the experience with the prototype.

\subsection{Brain-Computer Interface}\label{BCI}
The data obtained in INTENTION was used to train the BCI. For processing EEG and EMG data from INTENTION, we utilized the EEGLAB~\cite{Delorme2004-sn} toolbox with wrapper functions from BeMoBIL-pipeline~\cite{Klug2022-lc} running in the MATLAB (The MathWorks Inc. Natick, MA, USA) environment. First, to generate behavioral labels at a high temporal resolution for the EEG-based classifier, we leveraged EMG data from the flexor digitorum profundus. EMG amplitudes were band-pass filtered from 20 to 100 Hz and subsequently squared. Next, to label the time of movement onset, the EMG data was averaged across trials for the second preceding the screen tap. From this averaged data, the first sample where the EMG amplitude exceeded the 95th percentile was selected as the time of movement onset, see~\ref{fig:method}a.

% Ignore Next, noisy data segments were determined. 
% A segment was rejected as an outlier using Matlab's \textit{isoutlier} function using default parameters. % from isoutlier documentation (mean segment amplitude exceeding 3 scaled median absolute deviation)

Two event classes were then defined as follows: \textit{pre-movement} from -1000 to 0 ms preceding the (EMG detected) movement onsets and \textit{idle}, a one-second data segment between trials where participants were looking at a fixation cross.

\subsubsection{Preprocessing EEG and Selecting discriminative channels}\label{eeg_methods}
The EEG data was band pass filtered from 0.1 to 15 Hz. In the first step to prepare the EEG data for classifier training, noisy data segments were rejected. To this end, the EEGLAB function `autorej' was used, keeping default parameters. A trial was excluded if \textit{either} data from the \textit{idle} or \textit{pre-movement} class was rejected.

Then, to decrease the dimensionality of the EEG data, we selected discriminative channels following an approach outlined by~\citep{Schultze-Kraft2021-cu}. First, for both \textit{pre-movement} and \textit{idle} the mean signal in the last 100 ms of each segment was subtracted from the mean signal in the first 100 ms of the segment. The resulting value thus indicated how much the signal had changed in the course of the 1 s segments. In line with the literature, there should be a change over time in the \textit{pre-movement} segment but not in the \textit{idle} segment. Then, all channels were sorted (1) in descending order by the signal difference in the \textit{pre-movement} segments and (2) in ascending order by the signal difference in the \textit{idle} segments. With respect to the ability of the system to discriminate between movement intention and the absence of movement intention, an ideal channel should show a strong change in signal over the course of a \textit{pre-movement} segment but no difference during an \textit{idle} segment. The ranks of the two criteria were joined by summation and the resulting order was saved for later use in training and real-time application of the classifier. Lastly, channels C3, C4, and Cz were moved to the top of the order for every participant as these are most frequently reported in studies on the RP.

\subsubsection{Training of EEG classifier}
To classify EEG data, a linear discriminant analysis (LDA) with shrinkage regularization (automatic shrinkage using the Ledoit-Wolf lemma~\cite{Ledoit2004-bi}) was trained for each participant individually. As single-trial features, the (linear) slope coefficient was obtained for both \textit{idle} and \textit{pre-movement} segments by fitting a linear regression. The classifier was then trained using the slope coefficient of the selection of the most discriminative channels as features, generating a feature vector in the dimension of channels that were kept for classification, see~\ref{fig:method}b. We purposefully constrained the dimensionality of the feature vector to avoid over-fitting and decrease the computational load.

Using scikit-learn~\cite{Pedregosa2012-sj}, the LDA was cross-validated (using 5-folds) for a grid search from 6 to 20 channels with a step size of adding 2 channels. Following this grid search, the cross-validation with the highest accuracy determined the number of channels that were kept for training the model for real-time application. Ultimately, to determine the threshold at which, during the real-time application, the classifier would trigger the EMS, we computed the Receiver-operator characteristic (ROC) and from this selected the threshold at 15 \% false positive rate.

\subsubsection{Real-time application and EMS control}
During real-time application, the EEG data was buffered for the last second for the selected discriminative channels. The data was band-pass filtered analogously to the training data from 0.1 to 15 Hz. Next, the slope was computed. This procedure ran at an update rate of 10 Hz, hence every 100 ms a new prediction was obtained from the classifier. To smooth the prediction output to reduce false predictions due to unlikely peaks, the predicted probability for the \textit{pre-movement} class was smoothed by averaging the current and the preceding prediction with a weighting (weight of .3 for the preceding, and .5 for the current prediction). The weights were obtained through trial-and-error during piloting. Then, with a 10 Hz update rate, this smoothed probability, as well as the predicted class for the current frame were gating the EMS switch: when the probability exceeded the threshold and the currently predicted class was \textit{pre-movement}, the switch was opened for 0.5 s, see figure~\ref{fig:method}c.

% To trigger the EMS, an Arduino was used to flip a switch whenever triggering the attached EMS. We constrained the times when the switch could be flipped to specific moments in the experiment where participants were getting ready to perform an action, thereby limiting the impact of distracting EMS pulses at resting phases during the experimental trials. %This guaranteed that participants experienced EMS only during movement preparation

\subsubsection{Post-hoc evaluation of real-time performance}
To not break the user's focus on the task at hand, (subjective) labels were not obtained during the real-time application in AUGMENTED. In other words, participants were not directly queried, e.g. via questionnaires, to judge whether the stimulation in the current trial was in line with their intention or not. Without these labels, a post-hoc analyses of the binary classifiers performance was not possible, i.e. to ascertain for example false positive stimulations. However, we explored an alternate approach to estimate (subjective) labels that does not break the user's task immersion. We investigated prediction errors in response to the movement onset through event-related potentials (ERPs) at fronto-central electrode FCz. Previously, prediction error ERPs at, among others, electrode FCz have been shown to be one suitable candidate feature to detected breaks in (task) immersion, such as when perceiving glitches in VR~\cite{Gehrke2019-og, Gehrke2022-kz, Gehrke2024-xq, Terfurth2024-kh, Si-mohammed2020-ru}. In BCIs, these ERPs are frequently leveraged to correct system errors~\cite{Zander2016-ed}. In their work,~\citet{Zander2016-ed} demonstrated a BCI to decode a user’s intended cursor movement direction on a 6 × 6 grid. The system regularly probed the user by observing the EEG response to random cursor movements. How severely the random dot movement violated the user’s intention was directly reflected in the ERP activity. We hypothesized to find similar ERP signatures when the stimulation misaligned with the user's prediction.

Since this analyses was conducted post-hoc, it allowed for more signal processing in comparison to the feature extraction from INTENTION when participants were waiting to start the next block. With the goal to best recover the prediction error ERP, we again applied `BeMoBIL-pipeline'~\cite{Klug2022-lc} wrapper functions of EEGLAB~\cite{Delorme2004-sn}: After removing non-experiment segments at the beginning and end of the concatenated recording from all three conditions, EEG data was re-sampled to $250~Hz$. Next, bad channels were detected using the `FindNoisyChannel' function, which selects bad channels by amplitude, the signal-to-noise ratio, and correlation with other channels~\cite{Bigdely-Shamlo2015-ds}. Rejected channels were then interpolated while ignoring the EOG channel, and finally re-referenced to the average of all channels, including the original reference channel FCz. After applying a high-pass filter at 1.5 Hz, time-domain cleaning and outlier removal were performed using AMICA auto rejection~\cite{Palmer2011-zs}. Eye artifacts were removed using the ICLabel toolbox applied to the results from an AMICA~\cite{Pion-Tonachini2019-fy}. For this, ICLabel's popularity classifier was used, meaning that all components having the highest probability for the eye class were projected out of the sensor data.

For all three conditions, ERPs were extracted from band-pass filtered (0.1 Hz to 15 Hz) activity at electrode \textit{FCz}. These ERPs were obtained from -1000 ms to 500 ms around the movement onset. Trials were excluded in line with the removal for classifier training, see section~\ref{eeg_methods}. For all muscle-controlled trials without EMS, movement onset was defined as the time of the tap on the touchscreen minus the `EMG-delay' defined in section~\ref{BCI}. For all EMS-controlled trials, movement onset was defined as EMS onset. To keep only EMS-controlled trials where the stimulation resulted in an `immediate' screen tap, first all trials were rejected where there was no screen tap in the 350 ms following EMS. Next, `extreme outlier removal' was conducted on the delta between EMS and the screen tap using Tukey's method based on the inter-quartile range ~\cite{Tukey1949-sl}. In total, 32.9 (SD = 30.3) trials were rejected on average per participant.

% To analyze prediction error ERPs following the movement onset, potentially reflecting a disruptive experience impacting SoA, a linear mixed effects model was fit at each time point of the ERP using the `lme4' package~\cite{Bates2015-bh}. The model `ERP\_sample $\sim$ condition + (1|participantID)' was fit with condition reflecting trials belonging to INTENTION, INVOLUNTARY, or AUGMENTED. Prior to model fitting, all trials were aggregated per participant and condition. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `ERP\_sample $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. P-values were corrected for multiple comparisons using \textit{false discovery rate}~\cite{Benjamini1995-cw} at alpha = .05.

% Since the AUGMENTED condition contains both, `EMS-controlled' trials, where the classifier detected an intent to interact, as well as `muscle-controlled' trials, where the classifier failed to detect the user's intent and the user carried out the tap without EMS, we split these two trial categories. We hypothesized that `EMS-controlled' trials in AUGMENTED \textit{differ} from trials in INVOLUNTARY. The latter should elicit a stronger prediction error than the former since no impact of volition is present in INVOLUNTARY, which could potentially moderate the ERP. Similarly, we hypothesized that `muscle-controlled' trials in AUGMENTED should \textit{not} differ from INTENTION trials. To test these hypotheses, permutation t-tests were conducted on aggregated data using MNE-python~\cite{Gramfort2013-fa} contrasting `EMS-controlled' trials in AUGMENTED with INVOLUNTARY as well as `muscle-controlled' trials in AUGMENTED with INTENTION. The tests were conducted for the time window from 150-250 ms post movement-- or EMS-onset event. In short, these tests were conducted to investigate the moderating role of the user acting at their own volition in AUGMENTED.

\subsection{Measures of Agency Experience: Intentional Binding, Question \& Interview}% adapt title?
% To assess if SoA was preserved, an intentional binding measure, a standardized question, and an interview following the AUGMENTED block were investigated. 
After each condition participants were prompted to rate their experienced agency on a 7-point Likert scale with the statement ``It felt like I was in control of the movements during the task.'', the item was copied from~\cite{Bergstrom2022-fb}. Following AUGMENTED we interviewed users about their experience working with the system. After prompting users to recall their experience and summarize what their task had been, we set the focus to the tapping movement and asked them to ignore the time estimation task in the questions that followed. We entered the open part of the interview by asking: ``What did the system do?'' followed up by ``What was the difference between the three conditions''. After some time, and depending on their answers, we reset the focus to the AUGMENTED condition and asked ``How often was the system active?'' followed by ``What do you think caused the actions of the system?''. This was then followed up by an `open' interview in which we frequently asked `how' and `why' questions to inquire about the user's experience.

We analyzed the interviews by loosely following~\citet{Mayring2015-pp}. All interviews were manually transcribed and translated to English using DeepL\footnote{https://www.deepl.com/translator} (DeepL SE, Cologne, Germany). Before screening the texts, two experts clustered the responses into 3 clusters: First, ``Functionality'' referred to what participants attributed the source of the stimulation. Next, we clustered responses according to ``Guessed percentage'' of correct interaction, i.e., participants' estimate of how well the system was aligned with their intention. For the cluster ``Correct Interaction'', we specifically queried participants to recall the moments where the stimulation felt in line with their intention to move, then we clustered their responses into sentiments with positive and negative valence.

\subsection{Statistical Analyses}
To confirm and demonstrate the discriminative power of the EEG features, we plotted the amplitude time course of electrode Cz between \textit{pre-move} and \textit{idle} epochs. Electrode Cz was chosen for exemplary presentation, as it is frequently reported in studies on the RP, since it captures neural activity originating from the sensorimotor cortex. Next, the slope coefficients were extracted and a paired t-test was conducted. Ultimately, to assess the classifier performance, we calculated the F1 score, i.e. the harmonic mean of precision and recall, and plotted the ROC, see figure~\ref{fig:EEG_results}a to c.

\subsubsection{Hypotheses Testing}
Prior to any statistical analyses of the intentional binding measure, outlier trials were rejected. We applied `extreme outlier removal' using Tukey's method~\cite{Tukey1949-sl} on three time intervals that well describe `regular' behavior across trials: (1) Tapping the screen in a reasonable interval after the fixation cross disappeared. An excessively short or long delay indicated that participants either tapped the screen prematurely by accident or they were checking in with the experimenter, respectively. (2) Providing a `reasonable' estimation in the intentional binding task. (3) The EMS stimulation leading to an \textit{immediate} screen tap. A long delay between the EMS trigger and the subsequent screen tap indicated that the stimulation was not strong enough in this trial to lead to muscle actuation resulting in a screen tap. Taken together, applying Tukey's method to each of these time windows and fusing the rejected trials led to the exclusion of 110 trials across all participants (M = 12.2, SD = 9.1). % and a final data set of 2140 trials.

In line with the literature on intentional binding, we hypothesized that the time intervals should be underestimated for the INTENTION and the AUGMENTED conditions. This should not be the case for the INVOLUNTARY condition where users had no intention to move. Hence, when binding occurs, the intervals should be underestimated, hinting at a higher SoA.

To test this, we fitted a linear mixed effects model with \textit{condition} (INTENTION, INVOLUNTARY, AUGMENTED) as a fixed effect and `participantID' as a random effect on aggregated data. The model was specified as `time interval $\sim$ condition + (1|participantID)' and fit using the `lme4' package~\cite{Bates2015-bh}. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `time interval $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. We computed post-hoc pair-wise tests for `condition' corrected for multiple comparisons (Tukey method) using the emmeans package~\cite{Lenth2020-xk}

% We included all single trial data points in the analyses and did not aggregate before fitting the model as fitting the random effect `participantID' corrects for the correlation of the repeated measurements within participants.

Next, we hypothesized that subjective ratings of control over the tap movement are comparable between INTENTION and AUGMENTED and lower in the INVOLUNTARY condition. Again, we fitted a linear mixed effects model with \textit{condition} as a fixed effect and participant ID as a random effect. Coefficients were assessed in the same way as for the intentional binding parameter above.

In short, we tested two main hypotheses with regard to the agency experience in our three experimental conditions: (1) participants underestimate the tone delay when acting intentionally. Adding EMS in line with participants' intention to move does not affect this underestimation. (2) The subjective feeling of control is comparable between INTENTION and AUGMENTED conditions. INVOLUNTARY should decrease the feeling of control significantly. Additionally, we report the clustered interviews anecdotally.

To analyze prediction error ERPs following the movement onset, potentially reflecting a disruptive experience impacting SoA, a linear mixed effects model was fit at each time point of the ERP using the `lme4' package~\cite{Bates2015-bh}. The model `ERP\_sample $\sim$ condition + (1|participantID)' was fit with condition reflecting trials belonging to INTENTION, INVOLUNTARY, or AUGMENTED. Prior to model fitting, all trials were aggregated per participant and condition. A test statistic was obtained by calculating likelihood-ratio tests comparing the full model as specified above against the null model `ERP\_sample $\sim$ 1 + (1|participantID)'. All parameters were estimated by maximum likelihood estimation~\cite{Pinheiro2006-bk}. P-values were corrected for multiple comparisons using \textit{false discovery rate}~\cite{Benjamini1995-cw} at $\alpha = .05$.

Since the AUGMENTED condition contains both, `EMS-controlled' trials, where the classifier detected an intent to interact, as well as `muscle-controlled' trials, where the classifier failed to detect the user's intent and the user carried out the tap without EMS, we split these two trial categories. We hypothesized that `EMS-controlled' trials in AUGMENTED \textit{differ} from trials in INVOLUNTARY. The latter should elicit a stronger prediction error than the former since no impact of volition is present in INVOLUNTARY, which could potentially moderate the ERP. Similarly, we hypothesized that `muscle-controlled' trials in AUGMENTED should \textit{not} differ from INTENTION trials. To test these hypotheses, permutation t-tests were conducted on aggregated data using MNE-python~\cite{Gramfort2013-fa} contrasting `EMS-controlled' trials in AUGMENTED with INVOLUNTARY as well as `muscle-controlled' trials in AUGMENTED with INTENTION. The tests were conducted for the time window from 150-250 ms post movement-- or EMS-onset event. In short, these tests were conducted to investigate the moderating role of the user acting at their own volition in AUGMENTED.

%%% Resources
\begin{comment}

% An additional cluster ``takeaways'' was used to cluster other, miscellaneous,  interesting comments.

% "how did it feel like?". If user's made reference themselves to the EMS device we followed up by inquiring: "why do you think it (the EMS trigger) happened?", "What do you think triggered it?". The goal was to learn about the source attribution user's made about the EMS trigger. Since we leveraged the RP, we hypothesized that user's would attribute, at least in part, the EMS trigger to themselves, i.e. their movement intention. Next, we pressed deeper into learning about the cooperative nature between user and system. We asked "Do you believe you did it alone?" and when they made reference to a specific situation where the EMS was triggered we followed up with "What did it feel like, did you had the feeling the system controlled you or did you cooperate?" and "What would have to be there for it to feel like a cooperation?". We closed the interview with two additional exploratory questions: "Did the system influence your performance? If yes, how and why?" as well as asking them about any application scenarios they have in mind.

% \indent\textbf{(1) Intentional Binding.}
% At the end of each trial participants were tasked to estimate the delay between the screen tap and a subsequent tone. The real delay was pseudo-randomized out of [200, 350 and 500] ms. With 75 trials per condition, each real delay occurred 25 times.

% \indent\textbf{(2) Questionnaire.}

% \subsection{Brain-Computer Interface}
% The presented classification scheme required individual training data per participant, as transfer learning remains challenging with EEG~\cite{Wan2021-zz}. Therefore, we initially obtained training data on a variant of our task that did not include any muscle stimulation. In our task, participants were instructed, following an initial waiting period, to tap a touchscreen with their ring finger at their own volition, see figure ~\ref{fig:teaser}. 
% \subsubsection{Labelling movement classes using EMG}3

% Subsequently the data were labelled into two classes: an \textit{idle} class representing resting background activity and a \textit{pre-movement} class, representing the intent to (inter-)act.

% In summary, the experiment consisted of five phases: (1) a setup phase; (2) the BASELINE condition to obtain training data and labels; (3) the NO INTENTION condition: being passively moved by the EMS system; (4) the INTENTION EMS condition where the BCI was triggering the EMS system that in turn made participants tap the screen; and (5) an open interview to inquire about their experienced agency following the INTENTION EMS condition.

% \subsubsection{Experiment Design and Procedure}
% The first condition was used to obtain training data for the single-trial classification system. Here participants were wearing the EMG sensors instead of the EMS actuators and were instructed to tap the screen at their own volition (BASELINE, see instructions above). In the second condition, participants were instructed to hold and wait for the EMS system to move their finger thereby eliciting the screen tap (NO INTENTION). The timing for the EMS trigger was taken by randomly choosing a time between the 5th and 95th percentile of their actual reaction time in the BASELINE condition. In the third condition, the EMS trigger was controlled through the readiness potential classifier (INTENTION EMS). Participants were given the instruction: ``you will now work \textit{with} the system'' along the same instruction as in the BASELINE condition. The classifier was set to active after the fixation cross disappeared and until a screen tap was registered. Hence, the EMS was not triggered at other times during a trial, so as not to interfere with participants typing in their time estimation response.

% To not break the user's immersion by disturbing the flow of the interaction we decided against asking qualitative questions at the end of each and every trial.

% old:
% \subsubsection{Reproducing Results and Data Availability}
% Data, experimental protocol, analyses code including scripts for a reproduction of the presented results are hosted at open science foundation (OSF)\footnote{[anonymized]}. BIDS formatted data is hosted on openneuro~\cite{}.

% , and the interval between EMS stimulation and tap as fixed effects

% Hypothese:
% In line with the literature on intentional binding, we hypothesized this duration to remain stable in the case that agency was preserved during the test phase with EMS. A deviation of the tone duration from the training phase was hypothesized to reflect a disruption in experienced agency.

% operationalising temporal binding = underestimation (here or somewhere else)

% \indent\textbf{(1) Hand Tracking.} We used an HTC Vive Tracker, attached to the participant's wrist, to track their right hand at 90Hz. Therefore, we ran a simple Unity3D scene and streamed the data via the labstreaminglatyer (LSL)\footnote{https://github.com/sccn/labstreaminglayer}

% At the end of each trial participants were tasked to compare the duration of their arm movement to a tone and indicate whether the tone was longer or shorter. We set the initial duration at 500 ms for each participant. When participants indicated the tone to be longer or shorter by pressing either the up or down key we added, or respectively subtracted, 100 ms from the tone duration. This new duration was then used for the next trial and so on. With 60 trials in the training phase, we hypothesized participants would oscillate 50-50 around a target duration after about 2/3 of the training trials. The final duration after all training trials was then used again as the start duration during the test phase. In line with the literature on intentional binding, we hypothesized this duration to remain stable in the case that agency was preserved during the test phase with EMS. A deviation of the tone duration from the training phase was hypothesized to reflect a disruption in experienced agency.

% (1) participants waited for a new letter to appear on screen; (2) then, they were instructed to wait for a brief moment (~2s), before (3) reaching out and pressing the key. In line with the literature on the origin of the RP generating process they were told "to avoid preplanning the movement, avoid any obvious rhythm, and to press when they felt the spontaneous urge to move". (4) Two \todo{how many, but do keep this constant in order for them to focus on the duration of their action} seconds after they pressed the key, they were presented with a tone and asked to judge whether their movement, from the start of their reach to the button press, lasted longer or shorter than the tone. (5) When they perceived the tone to be longer, they had to press the up key or vice versa the down key for perceiving that the tone was shorter than their movement. Following an inter-trial-interval the next trial started.

% For each trial, a key was randomly selected from one of two groups of letters: (1) 'a', 's', 'd' or (2) 'j', 'k', 'l'. The selected group was then changed with every trial. This was done to keep the task engaging while maintaining a comparable reaching movement between trials.

% ERP with significance mask for Cz? Assessing significance of classifier

% IB Task: ttest of final ib time between with and without EMS -> # subjects input values
% did they increase the IB task in more than 50% of the trials after EMS?

% questionnaires: ttest against zero to report direction

% interviews: anecdotal summary and sentiment analyses with two groups splitting questionnaires at 0

% correlation with EEG classifier accuracy?

% evaluation of EEG classifier: can report false negatives, how often was it missing to predict a movement onset? hand was moving without EMS being triggered occurred how often? Not possible to evlaute false positives as no ground truth available

\end{comment}